---
title: "PRJEB5259 DP QUAL report" # replace this title accordingly
author: "jlai"
date: '2022-05-14' 
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
source("../utils/parser.R")
```

## Data input

Data input should be merged vcf files from the variant calling pipeline with snpeff annotations.

```{r load vcf file, results = F, message = F, echo = F}
in.vcf <- read_tsv("~/projects/dog_wes/PRJEB5259/results/merge_vcf/HaplotypeCaller_merge_snpEff.ann.vcf.gz", comment = "#", col_names = F) # replace input before hitting start/knit
```

## Parse out information

We are interested in the depth (DP) and quality of variant sites.

```{r parsing}
dp <- parse_info(in.vcf, "DP") %>% as.integer()
plot.df <- in.vcf %>% transmute(QUAL = X6, DP = dp)
```

## Visualize distributions

```{r plot1}
ggplot(plot.df,
       aes(x = DP)) +
  geom_density() +
  scale_x_continuous(trans = "log10") +
  theme_bw()

ggplot(plot.df,
       aes(x = QUAL)) +
  geom_density() +
  scale_x_continuous(trans = "log10") +
  theme_bw()
```

Observe that large majority of reads have very small DP. Let's see what the quantiles look like.

```{r quantiles}
quantile(plot.df$DP, seq(0,1,.1)) 
quantile(plot.df$QUAL, seq(0,1,.1))
```

The percentile of DP that resembles some useful depth to use (arbitrary, but we don't want something like 10 or so) is 90% percentile of the data.

Apply filter to retain the top 10% of DP.

```{r filter}
dp_cutoff <- quantile(plot.df$DP, seq(0,1,.1))[10] # starting at 0%, 90% is the 10th element
plot.df <- plot.df %>% 
  filter(DP > dp_cutoff)
```

```{r plot2}
ggplot(plot.df,
       aes(x = DP)) +
  geom_density() +
  scale_x_continuous(trans = "log10") +
  theme_bw()

ggplot(plot.df,
       aes(x = QUAL)) +
  geom_density() +
  scale_x_continuous(trans = "log10") +
  theme_bw()
```

With the DP filter, the QUAL metric is now shifted to the right (from the original data). Likely, we do not need to apply any further filters, as the minimum is \~30 (1 in 30 chance of a false result).
